---
title: "p8105_hw6_anw2158"
author: "Aung Nay Win"
output: github_document
---

```{r, include=FALSE}
library(dplyr)
library(broom)
library(purrr)
library(ggplot2)
library(tidyr)
library(boot)
library(modelr)
```


### Problem 1

```{r , warning=FALSE}
homicide_data = read.csv("./data/homicide-data.csv")

homicide_data = homicide_data |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    solved = ifelse(grepl("Closed by arrest", disposition, ignore.case = TRUE), 1, 0)
    ) |> 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black"),
         !is.na(victim_age)
         )
```

```{r}
baltimore_data = filter(homicide_data, city_state == "Baltimore, MD")

model_baltimore = glm(solved ~ victim_age + victim_sex + victim_race, 
                       data = baltimore_data, family = binomial())

tidy_model = tidy(model_baltimore)

tidy_model |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    odds_ratio = exp(estimate),
    lower_ci = exp(estimate - 1.96 * std.error),
    upper_ci = exp(estimate + 1.96 * std.error)
  )
```

```{r}
city_models = homicide_data |> 
  group_by(city_state) |> 
  do(
    tidy_glm = tidy(glm(solved ~ victim_age + victim_sex + victim_race, 
                        data = ., family = binomial()))
  )

city_or = city_models |> 
  unnest(tidy_glm) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    odds_ratio = exp(estimate),
    lower_ci = exp(estimate - 1.96 * std.error),
    upper_ci = exp(estimate + 1.96 * std.error)
  )
```

```{r}
city_or_ordered = city_or |> 
  arrange(odds_ratio)

ggplot(city_or_ordered, aes(x = reorder(city_state, odds_ratio), y = odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  coord_flip() +  
  labs(x = "City", y = "Adjusted Odds Ratio (Male vs Female Victims)", 
       title = "Adjusted Odds Ratios for Solving Homicides in Different Cities",
       subtitle = "Comparing Male Victims to Female Victims") +
  theme_minimal()
```
The plot shows the adjusted odds ratios (ORs) for solving homicides in various cities, comparing male to female victims. A significant variation in ORs across cities is observed, with some cities more likely to solve homicides involving male victims, while others show the opposite trend or no clear disparity. Points with confidence intervals not crossing the OR of 1 indicate statistically significant differences in the likelihood of solving cases based on victim gender. Notably, New York, NY, shows a much higher likelihood of solving homicides with female victims, whereas cities like Albuquerque, NM, display a higher likelihood when victims are male. The wide range of confidence intervals suggests varying degrees of certainty about these estimates, emphasizing the need to consider local context and data quality when interpreting these findings.

### Problem 2

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
boot_function = function(data, indices) {
  boot_data = data[indices, ]
  model = lm(tmax ~ tmin + prcp, data = boot_data)
  glance_data = glance(model)
  tidy_data = tidy(model)
  
  r_squared <- glance_data$r.squared
  
   beta1 <- tidy_data$estimate[tidy_data$term == "tmin"]
  beta2 <- tidy_data$estimate[tidy_data$term == "prcp"]
  if (beta1 <= 0 || beta2 <= 0) {
    log_betas <- NA  # Assign NA if product of betas is not positive
  } else {
    log_betas <- log(beta1 * beta2)
  }
  
  return(c(r_squared, log_betas))
}

set.seed(123)  
bootstrap_results = boot(data = weather_df, statistic = boot_function, R = 5000)
```

```{r}
r_squared_estimates <- bootstrap_results$t[,1]
log_betas_estimates <- bootstrap_results$t[,2]
```

```{r}
hist(r_squared_estimates, main = "Distribution of R^2 Estimates", xlab = "R^2")
```
The histogram provided represents the distribution of R^2 estimates obtained from 5000 bootstrap samples in a simple linear regression analysis with `tmax` as the response variable and `tmin` and `prcp` as predictors. The distribution appears to be fairly normal, centering around a high R^2 value, which suggests a strong explanatory power of the model. The concentration of estimates around the peak indicates consistency in the bootstrap resamples, implying that the model is stable across different samples drawn from the data. The shape of the distribution, without significant skewness or extreme outliers, provides confidence in the reliability of the R^2 statistic for this regression model in the given data set.

```{r}
hist(log_betas_estimates, main = "Distribution of log(beta1 * beta2) Estimates", xlab = "log(beta1 * beta2)")
```
The histogram shows the distribution of the estimated natural logarithm of the product of two regression coefficients (presumably beta 1 for tmin and beta 2 for prcp) from 5000 bootstrap samples. The distribution is skewed to the left, suggesting that the product of 
beta 1 and beta 2 is typically less than 1, since the logarithm of numbers less than 1 is negative. The tail extending towards the lower values indicates that there are occasional bootstrap samples where the product of the coefficients is very small. The concentration of the estimates around the mode of the distribution indicates that there's a common range for the logarithm of the product where most of the bootstrap samples fall, but the skewness signifies that the relationship captured by the model may vary across different samples, possibly due to the presence of outliers or influential points in the data.

```{r}
ci_r_squared <- quantile(r_squared_estimates, c(0.025, 0.975), na.rm = TRUE)
ci_log_betas <- quantile(log_betas_estimates, c(0.025, 0.975), na.rm = TRUE)
```

```{r}
list(ci_r_squared = ci_r_squared, ci_log_betas = ci_log_betas)
```

### Problem 3

```{r}
birthweight_df = read.csv("./data/birthweight.csv") |> 
  mutate(
    babysex = as.factor(babysex),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8, 9),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    malform = as.factor(malform)
  ) |> 
  mutate(
    frace = ifelse(frace == "Unknown", NA, frace),
    mrace = ifelse(mrace == "Unknown", NA, mrace)
  ) |> 
  na.omit()
```

Model 1: Hypothesized/Derived Model

Based on domain knowledge, factors like gestational age, mother's health (e.g., pre-pregnancy BMI, weight gain during pregnancy), and baby's health (e.g., presence of malformations) could be significant predictors.
```{r}
model_1 = lm(bwt ~ gaweeks + ppbmi + wtgain + malform, data = birthweight_df)
```

Model 2: Length and Gestational Age as Predictors
This model uses only length at birth and gestational age.
```{r}
model_2 = lm(bwt ~ blength + gaweeks, data = birthweight_df)
```

Model 3: Head Circumference, Length, Sex, and Interactions
This model includes all possible interactions between head circumference, length, and sex.
```{r}
model_3 = lm(bwt ~ bhead * blength * babysex, data = birthweight_df)
```

Cross-Validation and Prediction Error Comparison
We'll use cross-validation to compare models based on prediction error.
```{r}
cv_splits = modelr::crossv_mc(birthweight_df, 100)

cv_results = cv_splits |>  
  mutate(
    rmse_1 = map_dbl(seq_along(train), ~{
      train_set = train[[.x]]
      test_set = test[[.x]]
      model = lm(bwt ~ gaweeks + ppbmi + wtgain + malform, data = train_set)
      sqrt(mean((predict(model, newdata = test_set) - test_set$bwt)^2, na.rm = TRUE))
    }),
    rmse_2 = map_dbl(seq_along(train), ~{
      train_set = train[[.x]]
      test_set = test[[.x]]
      model = lm(bwt ~ blength + gaweeks, data = train_set)
      sqrt(mean((predict(model, newdata = test_set) - test_set$bwt)^2, na.rm = TRUE))
    }),
    rmse_3 = map_dbl(seq_along(train), ~{
      train_set = train[[.x]]
      test_set = test[[.x]]
      model = lm(bwt ~ bhead * blength * babysex, data = train_set)
      sqrt(mean((predict(model, newdata = test_set) - test_set$bwt)^2, na.rm = TRUE))
    })
  )
```

```{r}
avg_rmse = cv_results |> 
  summarise(
    avg_rmse_1 = mean(rmse_1),
    avg_rmse_2 = mean(rmse_2),
    avg_rmse_3 = mean(rmse_3)
  )

birthweight_df = birthweight_df |> 
  add_predictions(model_1) |> 
  add_residuals(model_1)

ggplot(birthweight_df, aes(x = pred, y = resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted values", y = "Residuals", title = "Residuals vs Fitted Plot for Model 1")
```

